---
title: "data.analysis.final.cuts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# Load Tidyverse, janitor and arcos, tidycensus, mapview, ggthemes, scales
library(tidyverse)
library(janitor)
library(arcos)
library(scales)
library(tidycensus)
#install.packages("ggthemes")
library(ggthemes)
#install.packages("mapview")
library(mapview)
library(ggrepel)
#install.packages('corrr')
library(corrr)
```

```{r}
# store one of our API keys as an object called key
key <- "uO4EK6I"
```

Loading crosswalk, which connects Census tracts and Continuums of Care.

```{r}
#import coc_crosswalk
coc_crosswalk <- read_csv("data/crosswalk-coc-to-tract.csv")
# API KEY
census_api_key("797ae2d903eb9bee9daa41bb71c0696897cbcba7")
# Define a list of states
states <- c("AL",	"AK",	"AZ",	"AR",	"CA",	"CO",	"CT", "DC", "DE",	"FL",	"GA",	"HI",	"ID",	"IL",	"IN",	"IA",	"KS",	"KY",	"LA",	"ME",	"MD",	"MA",	"MI",	"MN",	"MS",	"MO",	"MT",	"NE",	"NV",	"NH",	"NJ",	"NM",	"NY",	"NC",	"ND",	"OH",	"OK",	"OR",	"PA",	"RI",	"SC",	"SD",	"TN",	"TX",	"UT",	"VT",	"VA",	"WA",	"WV",	"WI",	"WY")
# Pull population from Census
population_2010 <- get_decennial(geography = "tract", variables = "P001001", state=states, year = 2010, geometry = FALSE)
population_2010 <- population_2010 %>%
  rename(total_2010_population = value)
coc_crosswalk <- coc_crosswalk %>%
  inner_join(population_2010, by=c("tract_fips" = "GEOID"))

```

Bringing in pills and population data for each Census tract in the country.
```{r}
#import tract_pills
tract_pills <- read_csv("data/tract_pills.csv") %>%
  select(-total_2010_population) %>%
  group_by(GEOID) %>%
  summarise(total_pills = sum(total_pills))
```

Combining Census tract pill data and CoCs.

```{r}

coc_crosswalk_x <- coc_crosswalk %>%
  inner_join(tract_pills, by=c("tract_fips" = "GEOID")) %>%
  group_by(coc_code, coc_name, state_code) %>%
  summarise(total_2010_population = sum(total_2010_population, na.rm = TRUE),
            total_pills = sum(total_pills, na.rm=TRUE))


```

Bringing in PIT count data, filtering for one year.

```{r}
pit_count <- read_csv("data/hud-pit-all.csv")

pit_count_2017 <- pit_count %>%
select(coc_code, coc_name, ends_with("_2017"))
```

Is there a correlation between unsheltered homelessness rate in 2017 and pills per capita?
```{r}
coc_population <- read_csv("data/coc_pills_2010.csv")

coc_crosswalk_x <- coc_crosswalk_x %>%
  inner_join(coc_population, by = "coc_code") %>%
select(coc_code, total_pills, total_2010_population.y) %>%
 mutate(pills_per_person = (total_pills/total_2010_population.y))

homelessness_rate_pills_per_person <- coc_crosswalk_x %>%
inner_join(pit_count_2017, by = "coc_code") %>%
select(coc_code, pills_per_person, unsheltered_homeless_2017, total_2010_population.y) %>%
mutate(unsheltered_homelessness_rate = (unsheltered_homeless_2017/total_2010_population.y)) %>%
arrange(desc(unsheltered_homelessness_rate))

```

```{r}
# Test correlation
homelessness_rate_pills_per_person %>%
  ungroup() %>%
  select(pills_per_person, unsheltered_homelessness_rate) %>%
  cor()

cor.test(homelessness_rate_pills_per_person$pills_per_person, homelessness_rate_pills_per_person$unsheltered_homelessness_rate)
```
The p value is less than .05, the correlation is statistically significant!!

Let's graph it:
```{r}
ggplot(homelessness_rate_pills_per_person) +
 geom_point(aes(log(unsheltered_homelessness_rate), pills_per_person)) +
 labs(x="Homelessness Rate", y="Pills per Person", title="As pain pills flowed in, unsheltered homelessness increased", caption = "Source: DEA ARCOS database, via Washington Post, HUD PIT Count data") +
 theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
 geom_smooth(aes(log(unsheltered_homelessness_rate), pills_per_person), method = "loess", se = FALSE) +
 geom_text(aes(log(unsheltered_homelessness_rate), pills_per_person, label=coc_code))
```

Filtering for Zillow Cluster to see if that impacts the correlation.
```{r}
#import coc_crosswalk
coc_crosswalk <- read_csv("data/crosswalk-coc-to-tract.csv")
# API KEY
census_api_key("797ae2d903eb9bee9daa41bb71c0696897cbcba7")
# Define a list of states
states <- c("AL",	"AK",	"AZ",	"AR",	"CA",	"CO",	"CT", "DC", "DE",	"FL",	"GA",	"HI",	"ID",	"IL",	"IN",	"IA",	"KS",	"KY",	"LA",	"ME",	"MD",	"MA",	"MI",	"MN",	"MS",	"MO",	"MT",	"NE",	"NV",	"NH",	"NJ",	"NM",	"NY",	"NC",	"ND",	"OH",	"OK",	"OR",	"PA",	"RI",	"SC",	"SD",	"TN",	"TX",	"UT",	"VT",	"VA",	"WA",	"WV",	"WI",	"WY")
# Pull
population_2010 <- get_decennial(geography = "tract", variables = "P001001", state=states, year = 2010, geometry = FALSE)
population_2010 <- population_2010 %>%
  rename(total_2010_population = value)
coc_crosswalk <- coc_crosswalk %>%
  inner_join(population_2010, by=c("tract_fips" = "GEOID")) %>%
filter(zillow_cluster_num == 1)
```

```{r}
#
coc_crosswalk_x <- coc_crosswalk %>%
  inner_join(tract_pills, by=c("tract_fips" = "GEOID")) %>%
  group_by(coc_code, coc_name, state_code) %>%
  summarise(total_2010_population = sum(total_2010_population, na.rm = TRUE),
            total_pills = sum(total_pills, na.rm=TRUE))


```

```{r}
coc_population <- read_csv("data/coc_pills_2010.csv")

coc_crosswalk_x <- coc_crosswalk_x %>%
  inner_join(coc_population, by = "coc_code") %>%
select(coc_code, total_pills, total_2010_population.y) %>%
 mutate(pills_per_person = (total_pills/total_2010_population.y))

homelessness_rate_pills_per_person <- coc_crosswalk_x %>%
inner_join(pit_count_2017, by = "coc_code") %>%
select(coc_code, pills_per_person, unsheltered_homeless_2017, total_2010_population.y) %>%
mutate(unsheltered_homelessness_rate = (unsheltered_homeless_2017/total_2010_population.y)) %>%
arrange(desc(unsheltered_homelessness_rate))

```
```{r}
# Test correlation
homelessness_rate_pills_per_person %>%
  ungroup() %>%
  select(pills_per_person, unsheltered_homelessness_rate) %>%
  cor()

cor.test(homelessness_rate_pills_per_person$pills_per_person, homelessness_rate_pills_per_person$unsheltered_homelessness_rate)
```
The correlation is even larger, and has a lower p value, so this is definitely something to look into.